Presentation Title

### 2017-08-20 thesis/chapters/ch3-model.tex
Edit everything

Our system consists of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
It is based on our prior work on TrustChain~\cite{multichain},
where every node independently interacts with each other via their own blockchain\footnote{It is originally called MultiChain but the name has changed to TrustChain}.

The goal of this chapter is to detail our system specification.
We begin our discussion with an intuitive overview of the architecture in \Cref{sec:system-overview}.
Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}.
Then, the three protocols which make up the complete system,
namely consensus protocol (\Cref{sec:cons-protocol}), transaction protocol (\Cref{sec:tx-protocol}) and validation protocol (\Cref{sec:vd-protocol}).
Finally, we discuss a few variations of our main design and their respective tradeoffs in \Cref{sec:tradeoffs}.

@@

\subsection*{TrustChain and Extended TrustChain}
Extended TrustChain naturally builds on top of the standard TrustChain. 
Thus we first describe the standard TrustChain.
Our description has minor differences compared to the description in~\cite{multichain},
which is to help with the description of the Extended TrustChain.
We remark the differences when they occur.

% \subsubsection*{Standard TrustChain}
In TrustChain, every node has a ``personal'' chain. 
Initially, the chain only contains a genesis block generated by the nodes themselves.
When a node $A$ wishes to add a new transaction (TX) with $B$,
a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
The TX block must have a valid hash pointer pointing to the previous block and a reference\footnote{This definition is different from the original TrustChain definition found in~\cite{multichain}.
    In the original definition, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
    This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
An example of is shown in \Cref{fig:trustchain-bad}.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{trustchain-bad}
    \centering
    \caption{
    A TrustChain example with three nodes.
    Every block is denoted by $t_{i,j}$, where $i$ is the node ID and $j$ is the sequence number of the block.
    The arrows represent hash pointers and the dotted lines represent references.
    The blocks at the ends of one dotted line are pairs of each other.
    The red block after $t_{b, 5}$ indicates a fork.}
    \label{fig:trustchain-bad}
\end{figure}

@@

% \subsubsection*{Extended TrustChain}
The primary difference is the introduction of a new type of block---checkpoint (CP) block.
The most important function of the CP block is to represent the state of a node, in the form of a hash pointer.
A collection of CP blocks from all nodes represents the state of the whole system.
Therefore, nodes need to reach consensus on CP blocks using some consensus protocol, which we explain next.
A visual representation is shown in \Cref{fig:trustchain-bad-cp}.

@@

The consensus protocol can be seen as a technique of running infinitely many times of some 
Byzantine consensus algorithm\footnote{More accurately it is asynchronous common subset (ACS) which we describe in \Cref{sec:acs-background}.},
starting a new execution immediately after the previous one is completed.
This approach is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
or CP blocks in our case.

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, at the beginning of every Byzantine consensus algorithm execution,
we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use them as the input to the Byzantine consensus algorithm, and then run it.
After running the algorithm,
the facilitators output a set of CP blocks which we call the consensus result.
The result is then propagated to the network and the next algorithm execution begins.

\subsection*{Transaction and validation}
The TX protocol is a simple request and response protocol.
The nodes exchange one round of messages and create new TX blocks on their respective chains.
Thus, as we mentioned before, one transaction should result in two TX blocks.

The consensus and transaction protocol by themselves do not provide a mechanism to detect malicious behaviour such as tampering the TrustChain.
Thus, we need a validation protocol to counteract such behaviour.
When a node wishes to validate one of its transactions, it asks the counterparty for the \emph{agreed fragment} of the transaction.
Which is a fragment of the counterparty's chain beginning and ending with CP blocks but contains the TX block belonging to that transaction,
the CP blocks must also be in consensus.
Upon the counterparty's response, the node checks whether the CP blocks are in some consensus result,
the hash pointers are valid and his TX is actually in the fragment.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because a cryptographically secure hash function is collision resistant)
to create a different chain that begins and ends with the same two CP blocks.

\subsection*{Complete protocol}
The complete protocol is essentially the concurrent composition of the three aforementioned protocols,
all making use the Extended TrustChain data structure.

@@

This section and the ones following it give a technical treatment of what the content in System overview.
For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) for most of this work.
\begin{itemize}
\item Lower case (e.g., $x$) denotes a scalar object or a tuple.
\item Upper case (e.g., $X$) denotes a set or a constant.
\item Sans serif (e.g., $\textsf{fn}(\cdot)$) denotes a function.
\item Monospace (e.g., $\texttt{ack}$) denotes message type.
\end{itemize}
Further, we use $a || b$ to denote the concatenation of the binary representations of $a$ and $b$.

@@

We assume there exists a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
Finally, we use the random oracle (RO) model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter.
Under the RO model, the probability of successfully computing the inverse of the hash function is negligible with respect to $\lambda$~\cite{bellare1993random}.

% TODO big N does not follow our convention
In our model, we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
$n \ge 3t + 1$ must hold.
This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network,
the exact reason becomes clear in~\Cref{sec:consensus-phase}.

@@

\section{Extended TrustChain}
The primary data structure used in our system is the Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h$ is the height of the block (i.e. $h = |B_u|$).
We often use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of all TX blocks in $B_u$ and $C_u$ is the set of all CP blocks is $B_u$,
then $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
% We assume there exists a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
% where $\tau$ represents the TX type and $\gamma$ represents the CP type.

@@

The TX block is a six-tuple, i.e
$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle.$$
We describe each item in turn.
\begin{enumerate}
\item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
\item $seq_u$ is the sequence number which should equal $i$.
\item $txid$ is the transaction identifier, it should be generated using a cryptographically secure pseudo-random number generator by the initiator of the transaction.
\item $pk_v$ is the public key of the counterparty $v$.
\item $m$ is the transaction message.
\item $sig_u$ is the signature created using $sk_u$ on the concatenation of the binary representation of the five items above.
\end{enumerate}
The fact that we have no constraint on the content of $m$ is in alignment with our design goal---application neutrality.

TX blocks come in pairs.
In particular, for every block 
$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
there exist one and only one \emph{pair} 
$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle.$$
Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
Thus, given a TX block, these properties allow us to identify its pair.
\end{definition}

@@

\begin{definition}
\textbf{\emph{Checkpoint block}}

The CP block is a five-tuple, i.e. 
$$c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle,$$
where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$,
the other items are the same as the TX block definition.
% Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.

@@

If the enclosure of some TX block $t_{u, i}$ is $\langle c_{u,a}, c_{u, b} \rangle$,
then its \emph{fragment} $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
For convenience, the function $\textsf{agreed\_fragment}(\cdot)$ represents the agreed fragment of some TX block if it exists, otherwise $\bot$.
\end{definition}

@@

Our consensus protocol $\Pi_c$ runs on top of Extended TrustChain.
It uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have an unbiased election of facilitators.
Concretely, we define the following.
\begin{definition}
\label{def:consensus}
\textbf{\emph{TrustChain consensus protocol}}

A TrustChain consensus protocol is correct if the following holds for every round $r$:
\begin{enumerate}
    \item \emph{Agreement}:
        If one correct node outputs a set of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties are similar to Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}), but there are subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds because a failure in one round will lead to failures in subsequent rounds.

@@

ACS has the nice property of censorship resilience when compared to other consensus algorithms.
For instance, Hyperledger and Tendermint use a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
In contrast, every party in ACS are involved in the proposal phase,
and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.

@@

The reader may notice from the system overview (\Cref{sec:system-overview}) that the facilitators are computed from the consensus result,
but the consensus result is agreed by the facilitators.
Thus we have a dependency cycle.
The goal of the bootstrap phase is to give us a starting point in the cycle.

@@

In practice, the bootstrap oracle is most likely the software developer and some of the desired properties cannot be achieved.
In particular, it is not possible to have the fairness property because it is unlikely that the developer knows the identity of every node in advance.

\subsection{Consensus phase}
\label{sec:consensus-phase}
The consensus phase begins when $\F_r$ is available to all the nodes.
Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS algorithm in round $r + 1$.
The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.

There are two scenarios in the consensus phase.
First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators,
where $c_{u, h}$ is the most recent CP block.
Second, if the node is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
% Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
Invalid messages are removed.
That is blocks with invalid signatures and blocks signed by the same key.
With the sufficient number of \texttt{cp\_msg} messages,
it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$.
We have to remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
which results in invalid blocks in the ACS output, but not in any of the inputs.

The core of the consensus phase is the ACS algorithm.
While any ACS algorithm that satisfies the standard definition will work,
we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS algorithm 
because it is a consensus algorithm designed for blockchain systems.
We do not use the full HoneyBadgerBFT due to the following.
First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transaction information.

Continuing, when $\F_{r}$ finish the ACS execution and reach agreement on $\C_{r+1}$,
they immediately broadcast two messages to all the nodes---
first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
The reason for sending \texttt{cons\_sig} is the following.
Recall that channels are not authenticated, 
and there are no signatures in $\C_{r + 1}$.
If a non-facilitator sees some $\C_{r + 1}$, it cannot immediately trust it because it may have been forged.
Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.

Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1})$ (\Cref{alg:new-cp}).
Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ (\Cref{alg:facilitator}),
and updates its facilitator set to the result.
This concludes the consensus phase and brings us back to the state at the beginning of the consensus phase,
so a new round can be started.

@@

\begin{algorithm}
\caption{Function $\textsf{get\_facilitator}(\C_r, n)$ takes the consensus result $\C_r$ and an integer $n$,
then sorts the CP blocks $C$ by the luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
\label{alg:facilitator}
\begin{algorithmic}
\State $\langle r, C \rangle \gets \C_r$
\State $\textsf{take} (n, \textsf{sort\_by} (\lambda x.\textsf{H}(\C_r || pk \text{ of } x), C))$
\end{algorithmic}
\end{algorithm}

@@

The TX protocol $\Pi_t$, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

\begin{algorithm}
    \caption{Function $\textsf{new\_tx}(pk_v, m, txid)$ generates a new TX block and appends it to the caller $u$'s chain.
    It is executed in the private context of $u$, i.e. it has access to the $sk_u$ and $B_u$.
    The necessary arguments are the public key of the counterparty $pk_v$, the transaction message $m$ and the transaction identifier $txid$.}
    \label{alg:new-tx}

@@

A key feature of the TX protocol is that it is non-blocking.
At no time in \Cref{alg:new-tx} or \Cref{alg:tx-proto} do we need to hold the chain state and wait for some message to be delivered before committing a new block to the chain.
This allows for a high level of concurrency where we can call $\textsf{new\_tx}(\cdot)$ multiple times without waiting for the corresponding \texttt{tx\_resp} messages.

@@

Up to this point, we do not provide a mechanism to detect tampering.
The validation protocol aims to solve this issue.
The protocol is also a request-response protocol, just like the transaction protocol.
But before explaining the protocol itself, we first define what it means for a transaction to be valid.

>>>>

Our system consists of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
It is based on our prior work on TrustChain~\cite{multichain},
where every node independently interacts with each other via their own hash chain\footnote{TrustChain is originally called MultiChain but the name has changed to TrustChain}.

The goal of this chapter is to detail our system specification.
We begin our discussion with an intuitive overview of the architecture in \Cref{sec:system-overview}.
Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}, 
and then the three protocols which make up the complete system.
Finally, we discuss a few variations of our design and their respective tradeoffs in \Cref{sec:tradeoffs}.

@@

\subsection*{TrustChain and Extended TrustChain}
Extended TrustChain naturally builds on top of the standard TrustChain. 
Thus we first describe the standard TrustChain.
Our description has a minor difference when compared to the description in~\cite{multichain},
we remark the difference when it occurs.

% \subsubsection*{Standard TrustChain}
In TrustChain, every node has a ``personal'' hash chain. 
Initially, the chain only contains a genesis block generated by the nodes themselves.
When node $A$ wishes to add a new transaction (TX) with $B$,
a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
The TX block must have a valid hash pointer pointing to the previous block and a reference\footnote{This definition is different from the original TrustChain definition found in~\cite{multichain}.
    In the original definition, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
    This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
An example of is shown in \Cref{fig:trustchain-bad}.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{trustchain-bad}
    \centering
    \caption{
    A TrustChain example with three nodes.
    Every block is denoted by $t_{i,j}$, where $i$ is the node ID and $j$ is the sequence number of the block.
    The arrows represent hash pointers and the dotted lines represent references.
    The blocks at the ends of the dotted lines are pairs of each other.
    The red block behind $t_{b, 6}$ indicates a fork.}
    \label{fig:trustchain-bad}
\end{figure}

@@

% \subsubsection*{Extended TrustChain}
Extended TrustChain introduces a new block type---checkpoint (CP) block.
The primary function of the CP block is to represent the state of a node, in the form of a hash pointer.
A collection of CP blocks from all nodes represents the state of the whole system.
Therefore, nodes need to reach consensus on CP blocks using some consensus protocol, which we explain next.
A visual representation is shown in \Cref{fig:trustchain-bad-cp}.

@@

The consensus protocol can be seen as a technique of running infinitely many times of a Byzantine consensus algorithm\footnote{More accurately it is an asynchronous common subset (ACS) algorithm, which we describe in \Cref{sec:acs-background}.},
starting a new execution immediately after the previous one is completed.
This approach is necessary because blockchain systems always need to reach consensus on new values proposed by the nodes in the system,
or CP blocks in our case.

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, at the beginning of every Byzantine consensus algorithm execution,
we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use those blocks as the input to the Byzantine consensus algorithm, and then run it.
After the algorithm completes,
the facilitators output a set of CP blocks which we call the consensus result.
The result is then propagated to the network and the next algorithm execution begins.

\subsection*{Transaction and validation protocols}
The TX protocol is a simple request and response protocol.
The nodes exchange one round of messages and create new TX blocks on their respective chains.
Thus, as we mentioned before, one transaction should result in two TX blocks.

The consensus and transaction protocol by themselves do not provide a mechanism to detect malicious behaviour such as tampering the TrustChain.
Thus, we need a validation protocol to counteract such behaviour.
When a node wishes to validate one of its transactions, it asks the counterparty for the \emph{agreed fragment} of the transaction.
Which is a fragment of the counterparty's chain beginning and ending with CP blocks but contains the TX block belonging to that transaction,
the CP blocks must also be in consensus.
Upon the counterparty's response,
the node checks whether the CP blocks are in some consensus result and some other conditions.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because a cryptographically secure hash function is collision resistant)
to create a different chain that begins and ends with the same two CP blocks.

\subsection*{The complete system}
The complete system is essentially the concurrent composition of the three aforementioned protocols,
all making use the Extended TrustChain data structure.

@@

This section and the ones following it give a technical explanation of \textsc{Checo}.
For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) for most of this work.
\begin{itemize}
\item Lower case (e.g., $x$) denotes a scalar object or a tuple.
\item Upper case (e.g., $X$) denotes a set or a constant.
\item Sans serif (e.g., $\textsf{fn}(\cdot)$) denotes a function.
\item Monospace (e.g., $\texttt{ack}$) denotes message type.
\end{itemize}
Further, we use $a || b$ to denote the concatenation of the binary representations of $a$ and $b$.

@@

Further, we assume there exists a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
Finally, we use the random oracle (RO) model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter.
Under the RO model, the probability of successfully computing the inverse of $\textsf{H}(\cdot)$ is negligible with respect to $\lambda$~\cite{bellare1993random}.

% TODO big N does not follow our convention
In our model, we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ nodes are malicious and the inequality
$n \ge 3t + 1$ must hold.
This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network to run the consensus protocol,
the exact reason becomes clear in~\Cref{sec:consensus-phase}.

@@

\section{Extended TrustChain}
The primary data structure used in our system is the Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h$ is the height of the block (i.e. $h = |B_u|$).
We use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of all TX blocks in $B_u$ and $C_u$ is the set of all CP blocks is $B_u$,
then $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
% We assume there exists a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
% where $\tau$ represents the TX type and $\gamma$ represents the CP type.

@@

The TX block is a six-tuple, i.e
$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle.$$
We describe each item in turn.
\begin{enumerate}
\item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
\item $seq_u$ is the sequence number, which should equal $i$.
\item $txid$ is the transaction identifier, it should be generated using a cryptographically secure pseudo-random number generator by the initiator of the transaction.
\item $pk_v$ is the public key of the counterparty $v$.
\item $m$ is the transaction message.
\item $sig_u$ is the signature created using $sk_u$ on the concatenation of the binary representation of the five items above.
\end{enumerate}
The fact that we have no constraint on the content of $m$ is in alignment with our design goal---application neutrality.

TX blocks come in pairs.
In particular, for every TX block 
$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
there exist one and only one \emph{pair} 
$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle,$$
if they are created by honest nodes running the TX protocol.
Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
Thus, given a TX block, these properties allow us to identify its pair.
\end{definition}

@@

\begin{definition}
\textbf{\emph{Checkpoint block and genesis block}}

The CP block is a five-tuple, i.e. 
$$c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle,$$
where $\C_r$ is the consensus result (which we describe next in \Cref{sec:consensus-result}) in round $r$,
the other items are the same as the TX block definition.
% Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.

@@

If the enclosure of some TX block $t_{u, i}$ is $\langle c_{u,a}, c_{u, b} \rangle$,
then its \emph{fragment} $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
For convenience, the function $\textsf{agreed\_fragment}(t_{u, i})$ outputs the agreed fragment of $t_{u, i}$ if it exists, otherwise $\bot$.
\end{definition}

@@

Our consensus protocol $\Pi_c$ runs on top of the Extended TrustChain.
It uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have an unbiased election of facilitators.
We formally define the desired properties below.
\begin{definition}
\label{def:consensus}
\textbf{\emph{TrustChain consensus protocol}}

A TrustChain consensus protocol is correct if the following holds for every round $r$.
\begin{enumerate}
    \item \emph{Agreement}:
        If one correct node outputs a set of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$\footnote{
                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}, and
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties are similar to the typical Byzantine consensus properties, but there are subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds because a failure in one round will lead to failures in subsequent rounds.

@@

ACS has the nice property of censorship resilience when compared to other consensus algorithms.
In PBFT~\cite{castro1999practical}, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
In contrast, every party in ACS are involved in the proposal phase,
and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.

@@

% The reader may notice from the system overview (\Cref{sec:system-overview}) that the facilitators are computed from the consensus result,
% but the consensus result is agreed by the facilitators.
% Thus we have a dependency cycle.
% The goal of the bootstrap phase is to give us a starting point in the cycle.

@@

In practice, the bootstrap oracle is most likely the software developer that sets up the system and assigns the facilitators of round 1 in a fair way.

\subsection{Consensus phase}
\label{sec:consensus-phase}
For any node $u$,
the consensus phase begins when $\F_r$ is available and the latest block is $c_{u, h}$.
Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS algorithm in round $r + 1$.
The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.

There are two scenarios in the consensus phase.
First, if $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
Second, if $u$ is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
% Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
Invalid messages are removed.
That is blocks with invalid signatures and blocks signed by the same key.
With the sufficient number of \texttt{cp\_msg} messages,
it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$.
We have to remove invalid blocks a second time because the adversary may send different CP blocks to different facilitators,
which results in invalid blocks in the ACS output, but not in any of the inputs.

The core of the consensus phase is the ACS algorithm.
While any algorithm that solves the ACS problem will work,
we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS algorithm 
because it is a consensus algorithm designed for blockchain systems.
We do not use the full HoneyBadgerBFT due to the following.
First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us.

Continuing, when $\F_{r}$ finish the ACS execution and reach agreement on $\C_{r+1}$,
they immediately broadcast two messages to all the nodes---
first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
The reason for sending \texttt{cons\_sig} is the following.
The channels are not authenticated, 
and there are no signatures in $\C_{r + 1}$.
If a non-facilitator sees some $\C_{r + 1}$, it cannot immediately trust it because it may have been forged.
Thus, to guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.

Upon receiving $\C_{r + 1}$ and at least $n - t$ valid signatures, $u$ performs two tasks.
First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1})$ (\Cref{alg:new-cp}).
Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ (\Cref{alg:facilitator}),
and updates its facilitator set to the result.
This concludes the consensus phase and brings us back to the state at the beginning of the consensus phase,
so a new round can be started.

@@

\begin{algorithm}
\caption{Function $\textsf{get\_facilitator}(\C_r, n)$ takes the consensus result $\C_r$ and an integer $n$,
then sorts the CP blocks $C$ by the luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
\label{alg:facilitator}
\begin{algorithmic}
\State $\langle r, C \rangle \gets \C_r$
\State \Return $\textsf{take} (n, \textsf{sort\_by} (\lambda x.\textsf{H}(\C_r || pk \text{ of } x), C))$
\end{algorithmic}
\end{algorithm}

@@

The TX protocol $\Pi_t$, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator then sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

\begin{algorithm}
    \caption{Function $\textsf{new\_tx}(pk_v, m, txid)$ generates a new TX block and appends it to the caller $u$'s chain.
    It is executed in the private context of $u$, i.e. it has access to the $sk_u$ and $B_u$.}
    \label{alg:new-tx}

@@

A key feature of the TX protocol is that it is non-blocking.
At no time in \Cref{alg:new-tx} or \Cref{alg:tx-proto} do we need to hold the chain state and wait for some message to be delivered before committing a new block to the chain.
This allows for a high level of concurrency where we can call $\textsf{new\_tx}(\cdot)$ and send \texttt{tx\_req} messages multiple times without waiting for the corresponding \texttt{tx\_resp} messages.

@@

The consensus protocol and the validation protocol do not provide a mechanism to detect tampering.
The validation protocol aims to solve this issue.
The protocol is also a request-response protocol, just like the transaction protocol.
It assumes that nodes store all the consensus results that are aware of.
But before explaining the protocol itself, we first define what it means for a transaction to be valid.

\subsection{Validity definition}
A transaction can be in one of three states---\emph{valid}, \emph{invalid} and \emph{unknown}.
Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
otherwise the node calling the function would have no point of reference.
This assumption is reasonable in practice because typically the caller is $u$,
so it knows its own TX block and the corresponding agreed fragment.
If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.
